---
title: "State of the Ecosystem:"
subtitle: "Reporting on Northeast Continental Shelf<br> ecosystem status and trends."
author: "Sean Hardison<br /> Ecosystem Dynamics and Assessment <br /> Northeast Fisheries Science Center"
output:
  xaringan::moon_reader:
    css: ["default", "libs/my-theme.css"]
    lib_dir: libs
    nature:
      titleSlideClass: ["right", "middle", "my-title"]
      highlightStyle: githubp
      highlightLines: true
      countIncrementalSlides: false
  revealjs::revealjs_presentation:
    self_contained: false
    reveal_plugins: ["notes", "search"]
---
class: top, left

```{r setup, include=FALSE}

options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(echo = F,
                      warning = F,
                      message = F)
#Plotting and data libraries
library(ggplot2)
library(dplyr)
library(tidyr)
library(ecodata)
library(here)
library(kableExtra)
library(ggrepel)
library(stringr)
library(patchwork)
library(grid)
library(plotly)
library(vegan)
library(rpart)
library(colorRamps)

#GIS libraries
library(sf)
library(rgdal)
library(raster)
library(rnaturalearth)

data.dir <- here::here("data")


#Time series constants
shade.alpha <- 0.3
shade.fill <- "lightgrey"
lwd <- 1
pcex <- 2
trend.alpha <- 0.5
trend.size <- 2
hline.size <- 1
hline.alpha <- 0.35
hline.lty <- "dashed"
label.size <- 5
hjust.label <- 1.5
letter_size <- 4
feeding.guilds <- c("Apex Predator","Piscivore","Planktivore","Benthivore","Benthos")
x.shade.min <- 2009
x.shade.max <- 2018
map.lwd <- 0.4
#Function for custom ggplot facet labels
label <- function(variable,value){
  return(facet_names[value])
}

#facet names for titles
facet_names <- list("Apex predators" = expression("Apex predators"),
                    "Piscivores" = expression("Piscivores"),
                    "Planktivores" = expression("Planktivores"),
                    "Benthivores" = expression("Benthivores"),
                    "Benthos" = expression("Benthos"))


#CRS
crs <- "+proj=longlat +lat_1=35 +lat_2=45 +lat_0=40 +lon_0=-77 +x_0=0 +y_0=0 +datum=NAD83 +no_defs +ellps=GRS80 +towgs84=0,0,0"

#Coastline shapefile
coast <- ne_countries(scale = 10,
                          continent = "North America",
                          returnclass = "sf") %>%
             sf::st_transform(crs = crs)

#State polygons
ne_states <- ne_states(country = "united states of america",
                                      returnclass = "sf") %>%
  sf::st_transform(crs = crs)

#high-res polygon of Maine
#new_england <- read_sf(gis.dir,"new_england")

#EPU shapefile
epu_sf <- ecodata::epu_sf %>% 
  filter(EPU %in% c("MAB","GB","GOM"))

```

---
## SOE 2019: Fish and invertebrates
**Objective: Keep harvested species within population ranges where productivity is maximized over long-term**
```{r NE-stock-status, fig.width = 8, fig.asp = 0.75, fig.align="center"}

#Get data, spread for plotting, and filter
stock_status <- ecodata::stock_status %>%
  spread(.,Var,Value) %>% 
  filter(Council %in% c("NEFMC","Both"))

#Plot constants
y.max <- 4
x.max <- 7.5

all_missing <- stock_status %>%
  filter(is.na(B.Bmsy),is.na(F.Fmsy)) %>% 
  dplyr::select(Code, Council)

b_missing <- stock_status %>%
  filter(is.na(B.Bmsy), !is.na(F.Fmsy)) %>% 
  dplyr::select(Code, Council)

f_missing <- stock_status %>%
  filter(is.na(F.Fmsy), !is.na(B.Bmsy)) %>% 
  dplyr::select(Code, Council)

#A dataframe that defines custom legend for stocks with unknown status

all.df <- data.frame(text = all_missing$Code,
                    x = rep(x.max,length(all_missing$Code)),
                    y = seq(4.35,2.85,-0.22)-0.5,
                    color = all_missing$Council)

b.df <- data.frame(text = b_missing$Code,
                    x = rep(x.max*0.8,length(b_missing$Code)),
                    y = c(4.35,4.14)-0.5,
                    color = b_missing$Council)

f.df <- data.frame(text = f_missing$Code,
                    x = rep(x.max*0.6,length(f_missing$Code)),
                    y = seq(4.35,2.85,-0.22)-0.5,
                    color = f_missing$Council)


#Plotting code
ggplot(data = stock_status) +
  geom_vline(xintercept = 1, linetype = "dotted", color = "grey60")+
  geom_vline(xintercept = 0.5, linetype = "dashed", color = "grey60")+
  geom_hline(yintercept = 1, linetype = "dashed", color = "grey60") +
  geom_point(aes(x = B.Bmsy,
                 y = F.Fmsy,
                 color = Council,
                 shape = Council)) +
  geom_text_repel(aes(x = B.Bmsy, #geom_text_repel auto-jitters text around points
                      y = F.Fmsy,
                      label = Code,
                      color = Council), show.legend = FALSE,nudge_y = 0.05, nudge_x = 0.05) +
  ylim(0,y.max) +
  xlim(0,x.max*1.1) +
  geom_text(data = all.df, aes(x = x, y = y, label = text, color = color),show.legend = FALSE, size = 3)+
  geom_text(data = b.df, aes(x = x, y = y, label = text, color = color),show.legend = FALSE, size = 3)+
  geom_text(data = f.df, aes(x = x, y = y, label = text, color = color),show.legend = FALSE, size = 3)+
  scale_color_manual(values = c("purple","blue"),#c("purple","blue"), #Change legend labels for clarity
                   name = "Managed by",
                   breaks = c("Both","NEFMC"),
                   labels = c("MAFMC/NEFMC","NEFMC"))+
  scale_shape_manual(values = c(1, 19), #Change legend labels for clarity
                     name = "Managed by",
                     breaks = c("Both","NEFMC"),
                     labels = c("MAFMC/NEFMC","NEFMC"))+
  annotate("text", x = 7.5, y = y.max, label = "F and B missing", fontface =2, size = 3)+
  annotate("text", x = 6, y = y.max, label = "B missing", fontface =2, size = 3)+
  annotate("text", x = 4.5, y = y.max, label = "F missing", fontface =2, size = 3)+
  xlab(expression(~B/B[msy])) +
  ylab(expression(~F/F[msy])) +
  ggtitle("New England Stock Status") +
  theme_ts()
```

* 4 stocks above F/F<sub>msy</sub> and below B/B<sub>msy</sub>
  * Southern NE Yellowtail, GOM Cod, CC-GOM Yellowtail, Southern GB/MAB red hake
  
???
Stocks depicted here represent a subset of the full ecosystem

---
## SOE 2019: Fish and invertebrates
**Objective: Keep harvested species within population ranges where productivity is maximized over long-term**

```{r stock-status, fig.width = 8, fig.asp = 0.75, fig.align="center"}

#Get data, spread for plotting, and filter
stock_status <- ecodata::stock_status %>%
  spread(.,Var,Value) %>% 
  filter(Council %in% c("MAFMC","Both"))

#Plot constants
y.max <- 2.0 #1.75 mackerel cut off F/Fmsy is 1.8
x.max <- 2.6

#A dataframe that defines custom legend for stocks with unknown status

unknown <- data.frame(text = c("Unknown Status", "Longfin",
                              "Illex", "N. Goosefish", "S. Goosefish"),
                    x = rep(0.9*x.max,5),
                    y = seq(0.93*y.max,1.4,-.1))


#Plotting code
ggplot(data = stock_status) +
  geom_vline(xintercept = 1, linetype = "dotted")+
  geom_vline(xintercept = 0.5, linetype = "dashed")+
  geom_hline(yintercept = 1, linetype = "dashed") +
  geom_point(aes(x = B.Bmsy,
                 y = F.Fmsy,
                 color = Council)) +
  geom_text_repel(aes(x = B.Bmsy, #geom_text_repel auto-jitters text around points
                      y = F.Fmsy,
                      label = Code,
                      color = Council), show.legend = FALSE,nudge_y = 0.05, nudge_x = 0.05) +
  scale_color_manual(values = c("purple","blue"), #Change legend labels for clarity
                     name = "Managed by",
                     breaks = c("Both","MAFMC"),
                     labels = c("MAFMC/NEFMC","MAFMC"))+
  ylim(0,y.max) +
  xlim(0,x.max) +
  geom_text(data = unknown, aes(x = x, y = y, label = text), #Custom legend for unknown stock status
            size = c(4.75,rep(4,4)),
            color = c("black",rep("blue",2), rep("purple",2))) +
  xlab(expression(~B/B[msy])) +
  ylab(expression(~F/F[msy])) +
  ggtitle("Mid-Atlantic Stock Status") +
  theme_ts()
```

* No stock above F/Fmsy and below B/Bmsy
* Mackerel currently seeing unsustainable fishing mortality

???
Plot shows the ratio of stock mortality (F) to stock mortality at maximum sustainable yield, where a number larger than one indicates the stock is experiencing unsustainable fishing pressure. On the x axis we have the ratio of estimated stock biomass to biomass at maximum sustainable yield. 