---
title: "State of the Ecosystem:"
subtitle: "Reporting on Northeast Continental Shelf<br> ecosystem status and trends."
author: "Sean Hardison<br /> Ecosystem Dynamics and Assessment <br /> Northeast Fisheries Science Center"
output:
  xaringan::moon_reader:
    css: ["default", "libs/my-theme.css"]
    lib_dir: libs
    nature:
      titleSlideClass: ["right", "middle", "my-title"]
      highlightStyle: githubp
      highlightLines: true
      countIncrementalSlides: false
  revealjs::revealjs_presentation:
    self_contained: false
    reveal_plugins: ["notes", "search"]
---
class: top, left

```{r setup, include=FALSE}

options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(echo = F,
                      warning = F,
                      message = F)
#Plotting and data libraries
library(ggplot2)
library(dplyr)
library(tidyr)
library(ecodata)
library(here)
library(kableExtra)
library(ggrepel)
library(stringr)
library(patchwork)
library(grid)
library(ggiraph)
library(vegan)
library(rpart)

#GIS libraries
library(sf)
library(rgdal)
library(raster)
library(rnaturalearth)

data.dir <- here::here("data")


#Time series constants
shade.alpha <- 0.3
shade.fill <- "lightgrey"
lwd <- 1
pcex <- 2
trend.alpha <- 0.5
trend.size <- 2
hline.size <- 1
hline.alpha <- 0.35
hline.lty <- "dashed"
label.size <- 5
hjust.label <- 1.5
letter_size <- 4
feeding.guilds <- c("Apex Predator","Piscivore","Planktivore","Benthivore","Benthos")
x.shade.min <- 2009
x.shade.max <- 2018
map.lwd <- 0.4
#Function for custom ggplot facet labels
label <- function(variable,value){
  return(facet_names[value])
}

#facet names for titles
facet_names <- list("Apex predators" = expression("Apex predators"),
                    "Piscivores" = expression("Piscivores"),
                    "Planktivores" = expression("Planktivores"),
                    "Benthivores" = expression("Benthivores"),
                    "Benthos" = expression("Benthos"))


#CRS
crs <- "+proj=longlat +lat_1=35 +lat_2=45 +lat_0=40 +lon_0=-77 +x_0=0 +y_0=0 +datum=NAD83 +no_defs +ellps=GRS80 +towgs84=0,0,0"

#Coastline shapefile
coast <- ne_countries(scale = 10,
                          continent = "North America",
                          returnclass = "sf") %>%
             sf::st_transform(crs = crs)

#State polygons
ne_states <- ne_states(country = "united states of america",
                                      returnclass = "sf") %>%
  sf::st_transform(crs = crs)

#high-res polygon of Maine
#new_england <- read_sf(gis.dir,"new_england")

#EPU shapefile
epu_sf <- ecodata::epu_sf %>% 
  filter(EPU %in% c("MAB","GB","GOM"))

```

---
## SOE 2019: Human dimensions
**Community engagement and reliance on fisheries**
.pull-left[
MAB: 

* Similar story in MAB, where there is heavy reliance on scallops

* Aquaculture in the MAB is also dominated by shellfish (oysters)

]
.pull-right[
```{r oyster-aqua, fig.height = 3}

aqua <- ecodata::aquaculture %>% 
  group_by(Var) %>% 
  mutate(hline = mean(Value)) %>%
  ungroup() %>% 
  mutate(Var = plyr::mapvalues(Var, from = c("md oyster harvest","nj oyster harvest","va oyster harvest"),
                                                    to  = c("MD","NJ","VA"))) %>%
  dplyr::rename(State = Var)

aqua$State <- factor(aqua$State, levels = c("VA","MD","NJ"))


aqua_plot <- ggplot() + 
  geom_segment(aes(x=2005,xend=2017,y=mean(aquaculture[aquaculture$Var == "va oyster harvest",]$Value),
                   yend=mean(aquaculture[aquaculture$Var == "va oyster harvest",]$Value)),
               size = hline.size,
           alpha = hline.alpha,
           linetype = hline.lty,
           color = "#1b9e77",
           inherit.aes = F) +
  geom_segment(aes(x=2012,xend=2016,y=mean(aquaculture[aquaculture$Var == "nj oyster harvest",]$Value),
                   yend=mean(aquaculture[aquaculture$Var == "nj oyster harvest",]$Value)),
               size = hline.size,
           alpha = hline.alpha,
           linetype = hline.lty,
           color = "#d95f02",
           inherit.aes = F) +
  geom_segment(aes(x=2012,xend=2017,y=mean(aquaculture[aquaculture$Var == "md oyster harvest",]$Value),
                   yend=mean(aquaculture[aquaculture$Var == "md oyster harvest",]$Value)),
               size = hline.size,
           alpha = hline.alpha,
           linetype = hline.lty,
           color = "#7570b3",
           inherit.aes = F) +
 #Highlight last ten years
  geom_line(data = aqua, aes(x = Time, y = Value, color = State), size = lwd) +
  geom_point(data = aqua,aes(x = Time, y = Value, color = State), size = pcex) +
  scale_color_manual(values = c(VA = "#1b9e77", MD = "#7570b3",NJ = "#d95f02")) +
  scale_x_continuous(breaks = seq(2005,2018,3),expand = c(0.01, 0.01)) +
  scale_y_continuous(labels = function(l){trans = l / 1000000})+
  ggtitle("Oyster harvest")+
  ylab(expression("Oysters sold (10"^6*" n)")) +
  xlab("")+
  theme_ts()

```

```{r com-eng2, fig.height = 9}
source("engagement_plots.R")
comm_plot <- engagement_plots(type = "comm",epu = "MAB")

aqua_plot + comm_plot + plot_layout(ncol = 1, heights = c(1, 3))
```
]

???
Virginia continues to lead the country in oyster production, with aquaculture in the region growing quickly. The time series looks flat compared to Virginia, but oyster aquaculture in New Jersey is also growing.

